---
title: "Assignment5"
author: "mutharasi"
date: "2024-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### 1) We will now try to predict per capita crime rate in the Boston data set.

a) Try out some of the regression methods explored in this chapter,such as best subset selection, the lasso, ridge regression, and PCR. Present and discuss results for the approaches that you consider.

```{r}
library(glmnet)
library(leaps)
library(pls)
library(MASS)
```
```{r}
boston=Boston
```


###### best subset selection

```{r}
fit.lm=lm(crim~.,data = boston)
summary(fit.lm)
```
In this model We zn,dis,rad,black,medv are having relationship with response variable.
```{r}
best_subset=regsubsets(crim~.,data = boston,nvmax = 13)
best_subset.summary<-summary(best_subset)
best_subset.summary$adjr2
```

```{r}
par(mfrow=c(1,3))
plot(best_subset.summary$cp,xlab = "variable",ylab="C_P",type = "l")
points(which.min(best_subset.summary$cp),best_subset.summary$cp[which.min(best_subset.summary$cp)],col="green",cex=3,pch=20)
plot(best_subset.summary$bic,xlab = "variable",ylab="BIC",type = "l")
points(which.min(best_subset.summary$bic),best_subset.summary$bic[which.min(best_subset.summary$bic)],col="green",cex=3,pch=20)
plot(best_subset.summary$adjr2,xlab = "variable",ylab="adjr2",type = "l")
points(which.max(best_subset.summary$adjr2),best_subset.summary$adjr2[which.max(best_subset.summary$adjr2)],col="green",cex=3,pch=20)
```


```{r}
coef(best_subset,which.min(best_subset.summary$cp))
```
###### ridge

```{r}
set.seed(2)
boston_matrix=model.matrix(crim~.,data = boston)[,-1]
```


```{r}
ridge=cv.glmnet(boston_matrix,boston$crim,alpha=0)
bestlam<-ridge$lambda.min
bestlam
```
```{r}
coef(ridge,s=bestlam)
```
###### lasso

```{r}
set.seed(1)
lasso=cv.glmnet(boston_matrix,boston$crim,alpha=1)
bestlam<-lasso$lambda.min
bestlam
```

```{r}
coef(lasso,s=bestlam)
```
##### pcr
```{r}
set.seed(2)
pcr=pcr(crim~.,data=boston,scale=TRUE,validation="CV")
pcr_summary<- summary(pcr)
pcr_summary
```

```{r}
validationplot(pcr,val.type = "MSEP")
```

```{r}
load_pcr<-pcr$loadings[,1:8]
load_pcr
```



b) Propose a model (or set of models) that seem to perform well on this data set, and justify your answer. Make sure that you areevaluating model performance using validation set error, crossvalidation, or some other reasonable alternative, as opposed to using training error.

```{r}
train_boston=sample(nrow(boston),nrow(boston)*0.70)
```
```{r}
tr_boston=boston[train_boston,]
```
```{r}
te_boston=boston[-train_boston,]
```

```{r}
set.seed(3)
tr_boston_matrix=model.matrix(crim~.,data = tr_boston)[,-1]
te_boston_matrix=model.matrix(crim~.,data = te_boston)[,-1]
```

```{r}
ridge=cv.glmnet(tr_boston_matrix,tr_boston$crim,alpha=0)
bestlam=ridge$lambda.min
bestlam
```

```{r}
pred_ridge=predict(ridge,s=bestlam,newx = te_boston_matrix)
test_error=mean((te_boston$crim- pred_ridge)^2)
test_error
```
```{r}
rmse= sqrt(test_error)
(rmse/mean(te_boston$crim))*100
```
```{r}
lasso=cv.glmnet(tr_boston_matrix,tr_boston$crim,alpha=1)
bestlam=lasso$lambda.min
```

```{r}
lasso_pred = predict(lasso, s = bestlam, newx = te_boston_matrix)
test_error = mean((te_boston$crim - lasso_pred)^2)
test_error
```

```{r}
rmse_lasso= sqrt(test_error)
(rmse_lasso/mean(te_boston$crim))*100
```

c) Does your chosen model involve all of the features in the data set? Why or why not?

The chosen model does not involve all of the features in the data set, because some  of the predictors are not statistically significant to response.
